---
title: "Homework #1, Mei Maddox."
author: "Mei Maddox   \n"
output:
  pdf_document: default
  html_document: default
---


**Submit the solution on Canvas into the corresponding assignment (e.g. "Homework #1") in the form of R Markdown report, knitted into either of the available formats (HTML, pdf or Word). Provide only code and output. NO NEED TO COPY THE PROBLEM FORMULATION (!)**


## Problem 1 (make sure to include code with outputs)

1. I will study the vegetarian variable, which takes on two categories: yes ("y") and no ("n"):

```{r, fig.height=3.5}
id <- "1IACzOQWcpvKqf3YnGye0qQOF2mtd42nK"
fl_data <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", id))
table(fl_data$vegetarian)
barplot(table(fl_data$vegetarian), ylab="Frequency", main="Student Vegetarianism")
```

The majority of students were non-vegetarians (51) whilst only 15% were vegetarian (9).

2. I will study the college gpa variable, which is a continuous variable on an interval from 0.0 to 4.0 inclusive (though the min for this sample is 2.6):

```{r, fig.height=3.5}
table(fl_data$college_GPA)
summary(fl_data$college_GPA)
barplot(table(fl_data$college_GPA), ylab="Frequency", xlab="GPA", main="Student GPA Breakdown")
hist(fl_data$college_GPA, xlab="GPA range", main="Student GPA Breakdown")
```

The distribution of the continuous college_gpa variable appears to be bimodal (with 3.0 and 3.5 being the modes) with a slight skew to the left (the mean is 0.047 smaller than the median with respect to a value range from [2.6, 4.0]. 

## Problem 2

**1.7** 

a) 3077 people who responded;

b) All adults in the United States;

c) 23%.

**1.8**

a) Sample: 1200 Floridians who participated in the poll;
   Population: All adult Floridians in the United States in 2006;

b) The poll does not survey the entire population, therefore the poll-reported percentages are quoted statistics of the sample which provide an estimate for the population parameter. 

**1.9** 

a) A graduate student;

b) A few students from a particular graduate program;

c) All students from that graduate program.

**1.10** 

a) Sample: The randomly selected citizens in the polled countries;
   Population: All citizens in the polled countries;

b) The poll does not survey the entire population, therefore the reported percentages are sample statistics, which provide an estimate for the population parameter.

**2.3**

a) Quantitative;

b) Categorical;

c) Categorical;

d) Quantitative.

**2.6**

a) Continuous;

b) Discrete;

c) Continuous;

d) Discrete.
