---
title: "Statistics Project"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(magrittr)

# Read in mass mobilization dataset
df <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSxIiUIhtrrcaJcrwcxzglHzG6st4dIVWiHtS-7lP6MBNp05U0fv2Sbo26ebCjLYPAnJqzAaaluS-XU/pub?gid=106237406&single=true&output=csv") %>%
  tibble::as_tibble()

# Column name alteration
col.names <- list(ccode='country_code', 
                  protestnumber='protest_number',
                  
                  # Dates
                  startday='start_day',
                  startmonth='start_month',
                  startyear='start_year',
                  endday='end_day',
                  endmonth='end_month',
                  endyear='end_year',
                  
                  # state responses
                  accomodation='accommodation',
                  
                  #protest information
                  protesterviolence='protester_violence',
                  protesteridentity='protester_identity',
                  
                  #protester_demands
                  price.increases..tax.policy='price_increases_or_tax_policy',
                  political.behavior..process='political_behavior_or_proces')

colnames(df) <- sapply(colnames(df), function(x){
  idx <- which(names(col.names)==x)
  if( length(idx) == 0 ) return( gsub('\\.', '_', x) )
  else return( col.names[idx] )
})


protester_demands_variables <- df %>% dplyr::select("labor_wage_dispute":"social_restrictions") %>% colnames()

state_responses_variables <- c('accommodation','ignore','arrests','crowd_dispersal','beatings','shootings','killings')

size.factors<-c('<50','50-99','100-999','1000-1999','2000-4999','5000-10000','>10000')

df %<>% 
  # Remove non-protests
  dplyr::filter(protest == 1) %>% 
  
  # Remove the protest column
  dplyr::select(-protest) %>%
  
  # Convert the following columns to factors
  dplyr::mutate_at(c("country", "region", "country_code", "protester_violence", "participants_category"), as.factor) %>%
  
  # Force protester demands and state response to be binary
  dplyr::mutate_at(20:31, function(x) as.integer(replace(x, x>1, 1))) %>%
  
  # Create date columns
  tidyr::unite("start_date", start_year:start_day, sep = '/', remove = F) %>% 
  tidyr::unite("end_date", end_year:end_day, sep = '/', remove = F) %>%
  dplyr::mutate_at(c("start_date", "end_date"), as.Date) %>%
  
  # Establish protest duration
  dplyr::mutate(duration=end_date-start_date+1)

# Rename levels 
levels(df$protester_violence) <- c("Non-Violent", "Violent")
#Add ordering to participants_category
df$participants_category %<>% ordered(levels=size.factors)

# Convert state responses to corresponding values or NA
for( idx in 1:length(state_responses_variables) ){
  df %<>% dplyr::mutate_at(state_responses_variables[idx], function(x) ifelse(x == 1, yes = idx - 2, no = NA))
}

# Create use of force metric
df %<>%  dplyr::mutate(use_of_force = (rowMeans(.[state_responses_variables], na.rm = T)+pmax(!!!rlang::syms(state_responses_variables), na.rm = T))/2)

# Create hemisphere metric
df %<>% dplyr::mutate(hemisphere = ifelse(grepl('America', .[["region"]]), yes = "west", no = "east"))
```

## Part 2


```{r}
ggplot2::ggplot(df, ggplot2::aes(use_of_force, hemisphere, fill= hemisphere)) + ggplot2::geom_boxplot() + ggplot2::ylab("Hemisphere") + ggplot2::xlab("Use of Force") + ggplot2::ggtitle("Use of Force by Hemisphere")


ggplot2::ggplot(df, ggplot2::aes(use_of_force, fill=hemisphere)) + ggplot2::geom_density(alpha=.2, size=1.05) + ggplot2::ylab("Density") + ggplot2::xlab("Use of Force") + ggplot2::ggtitle("Use of Force by Hemisphere")

t.test(df$use_of_force~df$hemisphere)
```


```{r}
# To check the validity of inference when sampling from such population, 
#    for VARIOUS SAMPLE SIZES, 
#  we proceed to:
#    - Treat the data as our population, calculate its mean mu_pop (our KNOWN population mean).
#    - Take samples of a fixed size from it, and test H0: mu = mu_pop for them (which we KNOW to be TRUE).
#    - Check how frequently H0 gets falsely rejected for 0.05 significance level.

ggplot2::ggplot(df, ggplot2::aes(use_of_force, fill= cut(use_of_force, 7))) + ggplot2::geom_histogram(bins=15) + ggplot2::ylab("Frequency") + ggplot2::xlab("") + ggplot2::ggtitle("Use of Force Raw Values")+ ggplot2::geom_vline(xintercept=mean(df$use_of_force, na.rm = T))

boxplot(df$use_of_force, horizontal = T)
title("Use of Force")


n.sim <- 10000
p.values <- numeric(n.sim)
ci.95 <- matrix(0, nrow=n.sim, ncol=2)
mu0 <- mean(df$use_of_force, na.rm = T)

samp.size <- 100
for (j in 1:n.sim){
  x <- sample(df$use_of_force, samp.size)
  p.values[j] <- t.test(x, mu=mu0)$p.value
  ci.95[j,] <- t.test(x, mu=mu0)$conf.int
}

# Type I error is generally a bit higher than expected (0.05).
mean(p.values < 0.05)
mean(ci.95[,1] < mu0 & ci.95[,2] > mu0)

ggplot2::ggplot() + ggplot2::aes(p.values, fill=p.values<0.05) + ggplot2::geom_histogram() + ggplot2::ylab("Frequency") + ggplot2::xlab("p-value") + ggplot2::ggtitle("Histogram of p-values") + ggplot2::geom_vline(xintercept=mean(p.values < 0.05), linetype="dashed", size=1) + ggplot2::annotate(x=.25,y=+Inf,label=paste("P( Type 1 Error) =",mean(p.values < 0.05)),vjust=3.5,geom="label")


ci.95 %>% tibble::as_tibble() %>% dplyr::mutate(mean=rowMeans(.)) %>% tibble::rowid_to_column() %>% ggplot2::ggplot(ggplot2::aes(x=mean, y=rowid, color=V1<mu0 & V2>mu0)) + ggplot2::geom_errorbar(ggplot2::aes(xmin=V1, xmax=V2)) + ggplot2::theme(axis.text.y=ggplot2::element_blank(),axis.ticks.y=ggplot2::element_blank()) + ggplot2::ylab("Trials") + ggplot2::xlab("Confidence Interval") + ggplot2::ggtitle("Use of Force Coverage") + ggplot2::geom_vline(xintercept=mean(mu0), linetype="dashed", size=1) + ggplot2::annotate(x=mu0,y=+Inf,label=paste("Coverage =",mean(ci.95[,1] < mu0 & ci.95[,2] > mu0)),vjust=3,geom="label") + ggplot2::scale_color_discrete(name = "Mu_0 Relative to CI", labels = c("Not Contained", "Contained"))
```



$\mu=\{average~use~of~force\}$
$H_0:\mu=0~~~vs.~~~H_a:\mu\ne0$
```{r}
x <- t.test(df$use_of_force, mu=2)

# TS
x$statistic

# p-value
x$p.value
```

1. $TS=$ `r x$statistic`
2. If $H_0$ were true, we would expect TS to take on values according to the t-distribution with `r x$parameter` degrees of freedom.
3. $p-value=$ `r `

Due to such a small p-value, have strong evidence to reject the null hypothesis and lean towards the alternative hypothesis (statistically significant)
```{r}
# Cohen d
abs(mean(df$use_of_force, na.rm = T) - 0)/sd(df$use_of_force, na.rm = T)
```
Large practical effect size, therefore has high practical significance

## Part 3

$\mu=\{average~use~of~force\}$
$H_0:\mu=0~~~vs.~~~H_a:\mu\ne0$
```{r}

ggplot2::ggplot(df, ggplot2::aes(protester_violence, fill=protester_violence=="Non-Violent")) + ggplot2::geom_bar() + ggplot2::ylab("Frequency") + ggplot2::xlab("") + ggplot2::ggtitle("Protester Violence")

x <- table(df$protester_violence) %>% binom.test(p=.7)

# TS
x$statistic

# p-value
x$p.value
```

1. $TS=$ `r x$statistic`
2. If $H_0$ were true, we would expect TS to take on values according to the t-distribution with `r x$parameter` degrees of freedom.
3. $p-value=$ `r `

Due to such a small p-value, have strong evidence to reject the null hypothesis and lean towards the alternative hypothesis (statistically significant)
```{r}
# Cohen d
abs(mean(df$use_of_force, na.rm = T) - 2)/sd(df$use_of_force, na.rm = T)
```
Medium practical effect size, therefore has moderate practical significance

Small practical effect size, therefore has low practical significance




```{r}
df %>% dplyr::select_if(is.numeric) %>% dplyr::select(-id, -accommodation) %>% dplyr::mutate(n_demands = rowSums(.[protester_demands_variables])) %>% replace(is.na(.), 0) %>% cor() %>% round(4) %>% highcharter::hchart(type="heatmap") %>%
highcharter::hc_plotOptions(heatmap=list(dataLabels=list(enabled=TRUE)))
```

