---
title: "Comprehensive Statistics Notes"
author: "Mei Maddox"
date: "Last Compiled on `r format(Sys.Date(), format='%b %d %Y')`"
description: |
  This is a minimal example of using the bookdown package to write a book.
  set in the _output.yml file.
  The HTML output format for this example is bookdown::gitbook,
output:
  bookdown::pdf_book: default
  bookdown::gitbook: default
  bookdown::epub_book: default
site: bookdown::bookdown_site
documentclass: book
bibliography: [packages.bib]
biblio-style: apalike
nocite: '@*'
link-citations: yes
github-repo: maddoxamei/Statistics
---



<!--chapter:end:index.Rmd-->

# Multiple Linear Regression: Extensions

<b style = "color:blue">Advantages</b> of classic linear regression models:

  - in many cases they <b style = "color:blue">work quite well</b> from <b style = "color:blue">prediction</b> standpoint,
  - they provide <b style = "color:blue">interpretable results</b>
  
On the other hand, they make some <b style = "color:red">restrictive assumptions</b>:

  - <b style = "color:red">Additivity</b>: The effect of changes in X on the response Y is constant, independent of what values the other predictors are held at; "Per 1-unit increase in X, <b style = "color:red">holding other predictors constant</b>, the average Y will change by <b style = "color:red">$\beta$ (constant)</b>"\

  - <b style = "color:red">Linearity</b>: Y constitutes a linear function of $X_1,X_2,...,X_p$\
  $Y\approx f(X_1,X_2,...,X_p)=\beta_0+\beta_1X_1+\beta_2X_2+\dots+\beta_pX_p$

## Interaction Effect: Removing the Additivity Assumption
Example: Predict the $Y = \{\#~of~units\}$ produced at a factory, based on

  - $X_1=\{\#~of~production~lines\}$
  - $X_2=\{\#~of~workers\}$
If <b style = "color:red">no workers</b> are available to operate the lines, increasing the # of lines will If <b style = "color:red">not be as effective</b> in increasing factory's production $\Longrightarrow$ there's an If <b style = "color:blue">interaction effect</b> between $X_1(\#~lines)$ and $X_1(\#~workers)$ in explaining $Y(\#~units~produced)$.

For 

[ISLR::Credit][ISLR::Credit] dataset

### Numerical Variables
$Balance = \beta_0+\beta_1Income+\beta_2Age+\beta_3Income\times Age$\
$Balance = \beta_0+(\beta_1+\beta_3Age)Income+\beta_2Age+\epsilon$\
$Balance = \beta_0+\tilde{\beta_1}Income+\beta_2Age+\epsilon$\

$\beta_3=${change in effect of Income ads per 1-unit increase in Age ads (or vice-versa)}



### Numerical and Categorical







$\color{red}{red~text}$
$\color{red}{\text{red text}}$


hence more workers, more effectivenes of lines lines or production

slide 7:
$H_0:\beta_3=0$\
$H_a:\beta_3\ne0$

slide 8: predictions on models with interaction
1) $\hat{sales}=6.75+0.019TV+0.029radio+0.001(TV\times radio)$
```{r eval=FALSE}
lm.obj <- lm(sales ~ TV + radio + TV:radio, data = Advertising)
lm.obj <- lm(sales ~ TV*radio, data = Advertising)
```

2) Interpret effect that 1000$ increase in TV budget has on sales

Per \$1000 increase in TV, holding radio constant, sales will increase by ($0.019+0.001\times radio$)$\times1000$ items, on average

3) Interpret effect that 1000$ increase in radio budget has on sales

Per \$1000 increase in radio, sales will increase by ($0.029+0.001\times TV$)$\times1000$ items, on average

slide 9:

1) Interpret the interaction coefficient $\hat{\beta_3}=0.0011$
Per \$1000 increase in TV ads, the effectiveness of radio ads will increase by $0.0011\times1000=1$ unit, on average

what is effectiveness?

2) Interpret coefficients $\hat{\beta_1}=0.0191$ and $\hat{\beta_2}=0.029$

For each \$1000 increase in TV ads, holding radio at \$0, sales will increase by 19 units, on average

slide 11:
$balance_i=\beta_0+$

limitations:
  - doesn't allow for different slopes between quantatitive predictor and response
  
same slope beta1hat means that the effect income has on credit balance does not depend on student status

limitation is addressed by adding interaction term $income\times D_{stud}$

$balance_i=\beta_0+\beta_1income_i+\beta_2D_{stud}+\beta_3(income_i\times D_{stud})+\epsilon_i,~\epsilon_i\sim_{ind}N(0,\sigma^2)$

$balance_i=\left\{ \begin{array}{cl}\beta_0+\beta_1income_i+\beta_2+\beta_3income_i+\epsilon_i \\ \beta_0+\beta_1income_i+\epsilon_i \end{array} \right.$

$balance_i=\left\{ \begin{array}{cl}(\beta_0+\beta_2)+(\beta_1+\beta_3)income_i+\epsilon_i,&i^{th}~person~(student) \\ \beta_0+\beta_1income_i+\epsilon_i,&i^{th}~person~(not~student) \end{array} \right.$

$H_0:\beta_3=0$
$H_a:\beta_3\ne0$
significant or no significant interaction???

<!--chapter:end:01-test.Rmd-->

# Appendix

## Dataset Documentation
### ISLR::Credit
```{r echo=FALSE, results='asis'}
help_console("Credit", "html", lines = -1, before = "<blockquote>", after = "</blockquote>")
```



<!--chapter:end:99-appendix.Rmd-->

